A finite-state morphological transducer for Kyrgyz

### 2012-03-14 paper/paper.tex
gone through todos

Nouns may be followed by possessive morphology before any case morphology.  This relates the noun to a preceding noun or pronoun in the genitive case.  However, when both possessive morphology and case morphology occur after a noun, there is some irregularity in the system.  Table \ref{irregposcase} summarises some of the forms.  Forms that do not result from simple concatenation of the possession and case endings are highlighted in bold as being irregular.

@@

%FIXME: is this about right?  and I mean all aspects of the paragraph
%FIXME: ZZ needs to be replaced with .. ?
To calculate the naïve coverage\footnote{Naïve coverage refers to the percentage of surface forms in a given corpora that receive at least one analysis.  Forms counted by this measure may have other analyses which are not delivered by the transducer.} of the analyser, two corpora were used.  The first corpus was a database dump of the Kyrgyz Wikipedia, dated 2011-09-23,\footnote{The exact name of the dump was \texttt{kywiki-20110923-pages-articles.xml}.} which was processed with the programs \texttt{aq-wikicrp} and ZZ to extract sentences. The second was a corpus generated from the archives of Radio Free Europe / Radio Liberty's Kyrgyz service, \href{http://www.azattyk.org}{azattyk.org}.  The RFERL corpus was built using a script that scraped the archives for all articles from 2010, which include articles on a wide variety of topics, from sports to politics and from culture to current events.

@@

%FIXME: * Some remaining transducer issues
Also, despite the good coverage, there are still a number of grammatical forms that have not been implemented into the transducer.  %FIXME: e.g.?
%these issues is fixed!
%Also, despite the good coverage, there are still a number of grammatical forms that have not been implemented into the transducer.  One of these is the irregular combination of 3rd person non-past finite verb plus the yes/no question particle.  For example, the third person non-past of /бар/ \eng{go} is [барат], but when the question particle -/BI/ is added, the resulting form is [барабы] instead of *[баратпы] (the transducer's current output).  There are also currently some problems handling the progressive aspect auxiliary /жат/ when it grammaticalises as a suffix with a couple verbs (/бар/ and /кел/) which often use a different participial form with this auxiliary than most verbs do.

>>>>

Nouns may be followed by possessive suffixes before any case suffixes.  This relates the noun to a preceding noun or pronoun in the genitive case.  However, when both possessive morphology and case morphology occur after a noun, there is some irregularity in the system.  Table \ref{irregposcase} summarises some of the forms.  Forms that do not result from simple concatenation of the possession and case endings are highlighted in bold as being irregular.

@@

%FIXME: is this about right?  and I mean all aspects of the paragraph
%FIXME: ZZ needs to be replaced with .. ?
To calculate the naïve coverage\footnote{Naïve coverage refers to the percentage of surface forms in a given corpora that receive at least one analysis.  Forms counted by this measure may have other analyses which are not delivered by the transducer.} of the analyser, two corpora were used.  The first corpus was a database dump of the Kyrgyz Wikipedia, dated 2011-09-23,\footnote{The exact name of the dump was \texttt{kywiki-20110923-pages-articles.xml}.} which was processed with the programs \texttt{aq-wikicrp} to extract sentences. The second was a corpus generated from the archives of Radio Free Europe / Radio Liberty's Kyrgyz service, \href{http://www.azattyk.org}{azattyk.org}.  The RFERL corpus was built using a script that scraped the archives for all articles from 2010, which include articles on a wide variety of topics, from sports to politics and from culture to current events.

@@

%FIXME: * Some remaining transducer issues
% maybe fixed now?
Also, despite the good coverage, there are still a number of grammatical forms that have not been implemented into the transducer.  For example, one such form is historically the negative of the eyewitness past tense, but is now used (mostly with a question clitic) in a certain type of modal expression;\footnote{A surprised contradictory expression, like ``Did I not just clean the floor??''} this cannot be tagged as the negative of the eyewitness past tense, since a suppletive form is tagged that way.  Other shortcomings of the transducer are that it is not set up to deal with affixes on abbreviations (e.g., АКШнын \eng{The USA's}) or capitalisation changes on lemmas that occur in different word classes (e.g., Финландия \eng{Finnland}, but финландиялык \eng{Finnish}).

%FIXME: e.g.?
%these issues is fixed!
%Also, despite the good coverage, there are still a number of grammatical forms that have not been implemented into the transducer.  One of these is the irregular combination of 3rd person non-past finite verb plus the yes/no question particle.  For example, the third person non-past of /бар/ \eng{go} is [барат], but when the question particle -/BI/ is added, the resulting form is [барабы] instead of *[баратпы] (the transducer's current output).  There are also currently some problems handling the progressive aspect auxiliary /жат/ when it grammaticalises as a suffix with a couple verbs (/бар/ and /кел/) which often use a different participial form with this auxiliary than most verbs do.
